Split-Apply-Combine with Dynamic Grouping

### 2024-03-27 paper/article.Rnw
Equation numbering, and abbreviations according to JSS style

The tables in Figure~\ref{fig:example} illustrate the idea. The left table
represents the data set to be aggregated by $A$ and $B$.  The table on the
right represents the output. Colors indicate which data was used. 
%
\begin{figure}[H]
\centering
\begin{tabular}{cccl}
\multicolumn{4}{l}{Input data}\\
\hline
$A$ & $B$ & $B_1$ & $Y$\\
\hline
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   1   \\
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   2   \\
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   3   \\
\cellcolor{green!25} 2 &                     12 & \cellcolor{green!25} 1 & \cellcolor{green!25} 4   \\
\cellcolor{green!25} 2 &                     12 & \cellcolor{green!25} 1 & \cellcolor{green!25} 5   \\
\cellcolor{green!25} 2 &                     13 & \cellcolor{green!25} 1 & \cellcolor{green!25} 6   \\
\cellcolor{blue!25}  3 &                     21 &                2       & \cellcolor{blue!25}  7   \\
\cellcolor{blue!25}  3 &                     22 &                2       & \cellcolor{blue!25}  8   \\
\cellcolor{blue!25}  3 &                     12 &                1       & \cellcolor{blue!25}  9   \\
\hline
\end{tabular}\hspace{1cm}\begin{tabular}{ccl}
\multicolumn{3}{l}{Output aggregates}\\
\hline
$A\times B$ & \code{level} & $\mu_Y$\\
\hline
\rowcolor{red!25}   1 11        & 0             & 2  \\
\rowcolor{green!25} 2 12        & 1             & 5  \\
\rowcolor{green!25} 2 13        & 1             & 5  \\
\rowcolor{blue!25}  3 21        & 2             & 8  \\
\rowcolor{blue!25}  3 22        & 2             & 8  \\
\rowcolor{blue!25}  3 12        & 2             & 8  \\
\hline
\end{tabular}
\label{fig:example}
\caption{Input data (left) with grouping variables $A$, $B$ and $B_1$, and
means of $Y$ per $A\times B$ after dynamic grouping (right).}
\end{figure}
%
The First row in the output represents group $(A=1,B=11)$. The collapsing level
is zero, which means that no collapsing was necessary. Indeed, in the data
table we see that there are three rows with $A=1$ and $B=11$ with $Y$ values
$1, 2$, and $3$, resulting in $\mu_Y=(1+2+3)/3=2$ for this group.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Balanced and Unbalanced Hierarchical Classifications}
Hierarchical classifications are abundant in (official) statistics.
They represent a classification of entities into non-overlapping, nested
groupings. Examples include the International Standard Industrial Classification
of Economic Activities (ISIC, \citet{un2022isic}), the related 
Statistical Classification of Economic Activities in Europe (NACE, \citet{eu2006nace})
and the Euroean Skills, Competences, Qualifications and Occupations classification
(ESCO, \citet{eu2022esco}).

>>>>

The tables in Figure~\ref{fig:example} illustrate the idea. The left table
represents the data set to be aggregated by $A$ and $B$.  The table on the
right represents the output. Colors indicate which data was used. 
%
\begin{figure}[H]
\centering
\begin{tabular}{cccl}
\multicolumn{4}{l}{Input Data}\\
\hline
$A$ & $B$ & $B_1$ & $Y$\\
\hline
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   1   \\
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   2   \\
\cellcolor{red!25}   1 &\cellcolor{red!25}   11 &                1       & \cellcolor{red!25}   3   \\
\cellcolor{green!25} 2 &                     12 & \cellcolor{green!25} 1 & \cellcolor{green!25} 4   \\
\cellcolor{green!25} 2 &                     12 & \cellcolor{green!25} 1 & \cellcolor{green!25} 5   \\
\cellcolor{green!25} 2 &                     13 & \cellcolor{green!25} 1 & \cellcolor{green!25} 6   \\
\cellcolor{blue!25}  3 &                     21 &                2       & \cellcolor{blue!25}  7   \\
\cellcolor{blue!25}  3 &                     22 &                2       & \cellcolor{blue!25}  8   \\
\cellcolor{blue!25}  3 &                     12 &                1       & \cellcolor{blue!25}  9   \\
\hline
\end{tabular}\hspace{1cm}\begin{tabular}{ccl}
\multicolumn{3}{l}{Output Aggregates}\\
\hline
$A\times B$ & \code{Level} & $\mu_Y$\\
\hline
\rowcolor{red!25}   1 11        & 0             & 2  \\
\rowcolor{green!25} 2 12        & 1             & 5  \\
\rowcolor{green!25} 2 13        & 1             & 5  \\
\rowcolor{blue!25}  3 21        & 2             & 8  \\
\rowcolor{blue!25}  3 22        & 2             & 8  \\
\rowcolor{blue!25}  3 12        & 2             & 8  \\
\hline
\end{tabular}
\label{fig:example}
\caption{Input data (left) with grouping variables $A$, $B$ and $B_1$, and
means of $Y$ per $A\times B$ after dynamic grouping (right).}
\end{figure}
%
The First row in the output represents group $(A=1,B=11)$. The collapsing level
is zero, which means that no collapsing was necessary. Indeed, in the data
table we see that there are three rows with $A=1$ and $B=11$ with $Y$ values
$1, 2$, and $3$, resulting in $\mu_Y=(1+2+3)/3=2$ for this group.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Balanced and Unbalanced Hierarchical Classifications}
Hierarchical classifications are abundant in (official) statistics.
They represent a classification of entities into non-overlapping, nested
groupings. Examples include the international standard industrial classification
of economic activities (ISIC, \citet{un2022isic}), the related 
statistical classification of economic activities in europe (NACE, \citet{eu2006nace})
and the Euroean skills, competences, qualifications and occupations classification
(ESCO, \citet{eu2022esco}).
